### 2020-08-17 article.tex
first init

\documentclass{article}

\usepackage[round]{natbib}
\usepackage{hyperref}
\usepackage{dirtree}
\usepackage{graphicx}
\graphicspath{{img/}}

\title{ARGoS-Blockchain interface}
\author{Volker Strobel \and Alex Pacheco \and Marco Dorigo}

\begin{document}
\maketitle

\section{Introduction}

Several recent works have addressed the combination of blockchain
technology and swarm robotics. This combinatin proves useful for
security, consensus finding, robot economies, robot-as-a-service, and
potentially many more.

However, running these experiments requires a suitable framework. Such
a framework proves useful both for researchers who do not have access
to a suitable robot platform, such as the Pi-puck robot. Additionally,
a suitable platform is important before running time- and
cost-consuming experiments with real robots.

The ARGoS robot simulator \citep{PinTriOGr-etal2012:si} is the
state-of-the-art research platform to conduct simulations in swarm
robotics.

In general, in swarm robotics research, each robot is a blockchain
node, and contributes to the maintenance of the system by exchanges
blockchain information. To link ARGoS and blockchain software, we
developed the ARGoS-Blockchain interface that provides access to the
blockchain nodes for the robots (Figure~\ref{fig:interface}). The
interface is intended to facilitate research in blockchain-based robot
swarms by allowing to call blockchain functions in
ARGoS. Additionally, Docker makes it easy to install and run the
interface on different platforms.

\section{Software availability}

%The Blockchain module is available at
%\url{https://github.com/Pold87/ARGoS-Blockchain-interface} and the
%ARGoS module is available at
%\url{https://github.com/Pold87/ARGoS-Blockchain-interface}.

%\dirtree{%
%.1 geth/.
%.1 img/.
%.1 local_scripts/.
%.1 .gitignore.
%.1 README.
%.1 docker-compose.yml.
%}

% TODO: this is a good figure, also include it
% \begin{figure}
%   \centering
%     \includegraphics[width=\textwidth]{Figure5}
%   \caption{For each robot, the ARGoS-Blockchain interface establishes
%     a connection to the Ethereum blockchain via a Docker container and
%     shell scripts that provide templates for executing Ethereum (geth)
%     functions.}
%   \label{fig:interface}  
% \end{figure}

\section{Overview}

The implementation of the custom Ethereum network is based on
Capgemini \textsc{aie}'s Ethereum
Docker\footnote{\url{https://github.com/Capgemini-AIE/ethereum-docker},
  accessed on November 6, 2019}. Docker containers
\citep{Mer2014:linux} contain all the necessary dependencies to run
specific applications and are more lightweight than a virtual
machine. In our setup, for each robot, the Ethereum implementation
\emph{geth} is executed in a separate Docker container. The simulated
robots maintain a \emph{custom} Ethereum network, i.e., a network that
is shared among the simulated robots and independent of Ethereum's
main network. Different containers can communicate with each other via
channels.

In order to execute an Ethereum function (e.g., create a new smart
contract) from ARGoS, a robot uses its C\texttt{++} controller to
attach to the Docker container. The Docker containers provide shell
scripts\footnote{The interface uses shell scripts, since, during
  development, it became evident that they are executed much faster
  than other Ethereum APIs.} with customizable templates (e.g., one of
the templates compiles the smart contract, uses the binary code to
send a blockchain transactions, and waits until the contract is
mined). Via Ethereum's \textsc{ipc} (interprocess communications)
interface, the shell scripts execute the Ethereum functions.

We use an auxiliary `bootstrap' node for publishing the smart contract
to the blockchain at the beginning of each run of the simulations
(Figure~\ref{fig:initialization}). The bootstrap node then mines the
smart contract and sends the contract address and the \textsc{abi}
(application binary interface; the \textsc{abi} specifies which
functions a smart contract provides and how to call them) to the
controllers of the robots. As soon as this is done, the bootstrap node
is removed from the network. The bootstrap node is not necessarily
required and the smart contract could also be created by a
robot. However, we used an auxiliary node to make sure (i) that the
smart contract is available at the start of the actual experimental
run and (ii) that robots have the same initial conditions in all
experiments.

The experiments were conducted on a computer cluster. To simulate the
limited hardware of real robots, one core with 2.0~GHz and 1.8~GB of
memory was assigned to each Docker container\footnote{This is a
  reasonable choice as a robot's computer could easily have such
  characteristics. It is also a convenient choice because on a
  computer with 2.0~GHz and 1.8~GB of RAM, Ethereum works
  ``out-of-the-box,'' without any modifications; therefore, any
  interested user can obtain the most recent release of Ethereum from
  the official depository and use it with our publicly available
  ARGoS-Blockchain interface.}. The communication channels between the
Docker containers were only established when robots were within a
50~cm communication range in order to simulate the local communication
capabilities of real robots.

The description in this report is intended to allow everyone to run
blockchain-based robot experiments in simulation. There is no need for
specific hardware but powerful CPU and RAM are advantageous for fluent
experiments.

The ARGoS-Blockchain interface is composed of two modules.

\subsubsection{Module~1: Blockchain part}
Module~1 allows for creating an Ethereum network with several nodes,
where each node is located in a separate Docker container. For the
ARGoS-Blockchain interface, the interaction with the Ethereum nodes is
done via C\texttt{++}, using the code in the following repository:
\url{https://github.com/Pold87/robot-swarms-need-blockchain}

This repository contains one module of the ARGoS-Blockchain interface
that is described in the article ``Blockchain Technology Secures Robot
Swarms: A Comparison of Consensus Protocols and Their Resilience to
Byzantine Robots by Strobel, V., Castello Ferrer, E., and Dorigo, M.''

For debugging purposes or for creating your own private Ethereum
network, you can also use this module without ARGoS.

\begin{figure}[t]
  \centering
  \includegraphics{interface}
  \caption[Overview of the ARGoS-Blockchain interface]{This scheme
    gives an overview of the workflow of the ARGoS-Blockchain
    interface}
  \label{fig:interface}
\end{figure}

\subsubsection{Module~2: ARGoS part}

\section{Setup}

The very first time you run this code, it is required to create the
Docker image for the Ethereum nodes and initialize Docker Swarm as
follows:

\begin{verbatim}
cd geth/
docker build -t mygeth .
docker swarm init
\end{verbatim}

Additionally, you have to set the variable \verb|DOCKERFOLDER| in the
file \verb|global_config.sh| to the full path where this repository is
located on your computer, for example:

\begin{verbatim}
/Users/vstrobel/Documents/ARGoS-Blockchain-interface/
\end{verbatim}

In order to be able to mine, you need to create the DAG datasets as
follows (the creates files require approximately 2~GB disk space and
the execution of the script can take take several minutes):

\begin{verbatim}
cd local_scripts/
bash create_dag.sh
\end{verbatim}

\section{Run}
Usually, the network is created when a swarm robotics experiment is
started, using one of the start scripts in
\url{https://github.com/Pold87/robot-swarms-need-blockchain}.

However, you can also start the Ethereum network without ARGoS, using
the following command:

\begin{verbatim}
bash local_scripts/start_network.sh <number of nodes>
\end{verbatim}

That is, \verb|bash local_scripts/start_network.sh 5|, would create a
private Ethereum network with 5 nodes.

\subsubsection*{Acknowledgements}

Volker Strobel and Marco Dorigo acknowledge support from the Belgian
F.R.S.-FNRS, of which they are a Research Fellow and a Research
Director respectively. Alexandre Pacheco acknowledges support from the
Faculty of Applied Sciences of the Universit\'{e} Libre de Bruxelles.

%\bibliographystyle{splncs04}
\bibliography{library}

\end{document}